---
title: "Vulcanoplot"
author: "Anja Bråthen Kristoffersen"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r include = F}
knitr::kable(paste0("date: ", Sys.Date()))

library(reshape2)
library(ggplot2)
library(psych)
library(ggrepel)
library(censReg)

utSti <- fs::path("F:", "Forskningsprosjekter", "PDB 2794 - Immune responses aga_", "Forskningsfiler", "JOBO", "CyTOF", "Analyse i R OUS", "CleanUpGatingMarch2022", "Analyse", "Endelig", "Figurer")

dataSti <- "F:/Forskningsprosjekter/PDB 2794 - Immune responses aga_/Forskningsfiler/Datafiler/"
#utSti <- "F:/Forskningsprosjekter/PDB 2794 - Immune responses aga_/Forskningsfiler/ABRK/FigurerKorrelasjon/"
dat1 <- read.csv2("F:/Forskningsprosjekter/PDB 2794 - Immune responses aga_/Forskningsfiler/Datafiler/MergedDataT1.csv")
dat2 <- read.csv2("F:/Forskningsprosjekter/PDB 2794 - Immune responses aga_/Forskningsfiler/Datafiler/MergedDataT2.csv")
dKont <- read.csv2("F:/Forskningsprosjekter/PDB 2794 - Immune responses aga_/Forskningsfiler/Datafiler/NederlandKOntroll.csv")


gener <- colnames(dat1)[which(colnames(dat1) == "AIRE"): which(colnames(dat1) == "ZNF532")]
# dGenGrupper <- as.data.frame(readxl::read_excel("F:/Forskningsprosjekter/PDB 2794 - Immune responses aga_/Forskningsfiler/Data fra Nederland April 2020/Table_List of target genes og refs for dcRT-MLPA.xlsx", skip = 2))

#length(dat2$Pasientnr[dat2$Pasientnr %in% dat1$Pasientnr])

# dGenGrupper$Gene[which(!dGenGrupper$Gene %in% colnames(dat1))]
# #disse må endres navn på
# colnames(dat1)[grep("PTPRC", colnames(dat1))]
# colnames(dat1)[grep("IL4", colnames(dat1))]
# colnames(dat1)[grep("CD274", colnames(dat1))]
# colnames(dat1)[grep("IFIT", colnames(dat1))]
# 
# dGenGrupper$Gene[dGenGrupper$Gene == "PTPRCv1/CD45RA"] <- "PTPRCv1"
# dGenGrupper$Gene[dGenGrupper$Gene == "PTPRCv2/CD45RO"] <- "PTPRCv2"
# dGenGrupper$Gene[grep("IL4", dGenGrupper$Gene)[2]] <- "IL4d2"
# dGenGrupper$Gene[dGenGrupper$Gene == "CD274/PD-L1"] <- "CD274"
# dGenGrupper$Gene[dGenGrupper$Gene == "IFITM1/3"] <- "IFITM1.3"
# 
# dGenGrupper$Gene[which(!dGenGrupper$Gene %in% colnames(dat1))]
# 
# for(i in 1:nrow(dGenGrupper)){
#   if(is.na(dGenGrupper$Module[i])){
#     dGenGrupper$Module[i] <- dGenGrupper$Module[i-1]
#   }
# }

# utvalgteGrupper <- c("Immune cell subset markers - B cells", "T cell subset markers", "Th1 associated genes", "Th2 associated genes", "Th17 associated genes", "Treg associated genes", "Cytotoxicity markers", "Myeloid associated genes", "IFN signaling genes" )
# 
# dGenGrupper$utvalgteGener <- dGenGrupper$Module %in% utvalgteGrupper
# nUtvalgt <- sum(dGenGrupper$utvalgteGener)
# 
# 
# dGenGrupper <- dGenGrupper[1:144,] #tar ut householder gene
# gener <- dGenGrupper$Gene[!grepl("Housekeeping", dGenGrupper$Module)]
# 
# 
# dat1$crbny <- dat1$Mental + I(dat1$Resp > 29) + I(dat1$Syst.BT < 90) + I(dat1$Diast.BT < 61)
# dat1$crb.alvor <- dat1$crbny > 0
# #table(dat1$crb.alvor)


dat1$Dager.innlagtFlerEnn4 <- 0
dat1$Dager.innlagtFlerEnn4[dat1$`Dager innlagt` > 4] <- 1
dat1$OxPnem5D <- apply(dat1[,c("Ox", "Pneumoniinfilt", "Dager.innlagtFlerEnn4")], 1, sum)

dat1$alvorligSykModSept2020 <- 0
dat1$alvorligSykModSept2020[dat1$OxPnem5D == 3 ] <- 4
dat1$alvorligSykModSept2020[dat1$OxPnem5D == 2 ] <- 3
dat1$alvorligSykModSept2020[dat1$OxPnem5D == 1 ] <- 2
dat1$alvorligSykModSept2020[dat1$OxPnem5D == 0 ] <- 1
dat1$alvorligSykModSept2020[dat1$Intensiv == 1] <- 5  #1 = ja, 2 = nei

dat1$Status <- "Moderate"
dat1$Status[dat1$alvorligSykModSept2020 > 2] <- "Severe"





dat2$crbny <- dat2$Mental + I(dat2$Resp > 29) + I(dat2$Syst.BT < 90) + I(dat2$Diast.BT < 61)
dat2$crb.alvor <- dat2$crbny > 0
#table(dat1$crb.alvor)

dat2$Dager.innlagtFlerEnn4 <- 0
dat2$Dager.innlagtFlerEnn4[dat2$`Dager innlagt` > 4] <- 1
dat2$OxPnem5D <- apply(dat2[,c("Ox", "Pneumoniinfilt", "Dager.innlagtFlerEnn4")], 1, sum)

dat2$alvorligSykModSept2020 <- 0
dat2$alvorligSykModSept2020[dat2$OxPnem5D == 3 ] <- 4
dat2$alvorligSykModSept2020[dat2$OxPnem5D == 2 ] <- 3
dat2$alvorligSykModSept2020[dat2$OxPnem5D == 1 ] <- 2
dat2$alvorligSykModSept2020[dat2$OxPnem5D == 0 ] <- 1
dat2$alvorligSykModSept2020[dat2$Intensiv == 1] <- 5  #1 = ja, 2 = nei

dat2$Status <- "Moderate"
dat2$Status[dat2$alvorligSykModSept2020 > 2] <- "Severe"



dat1$bakterieFunn <- !is.na(dat1$Bakt1) | !is.na(dat1$Bakt2)
dat2$bakterieFunn <- !is.na(dat2$Bakt1) | !is.na(dat2$Bakt2)
#dat1$posluftvei <- ( dat1$Ekspektorat == 1 )| ( dat1$Nasopharynx == 1)
```


```{r, echo = F, fig.width=5, fig.height=5}
pVerdiStjerne <- function(p){
  if(p < 0.001){
    return("***")
  } else {
    if(p < 0.01){
      return("**")
    } else { 
      if(p < 0.05){
        return("*")
      } else {
        return("-")
      }
    }
  }
}


plotPerGenT1T2 <- function(x, y, z, tittel = "", miny, maxy){
  pverdixy <- pVerdiStjerne(wilcox.test(x,y, paired = T)$p.value)
  pverdixz <- pVerdiStjerne(wilcox.test(x,z)$p.value)
  pverdiyz <- pVerdiStjerne(wilcox.test(y,z)$p.value)
  
  boxplot(x,y,z, label = c("T1", "T2", "Friske"), ylim = c(miny, maxy), main = tittel)
  points(rep(1, length(x)), x)
  axis(2)
  axis(1, at = c(1,2,3), label = c("T1", "T2", "Friske"))
  points(rep(2, length(y)),y)
  for(i in 1:length(x)){
    lines(c(1,2),c(x[i], y[i]), col = ((x[i] - y[i]) > 0) + 2) 
  }
  points(rep(3, length(z)), z)
  legend("topleft", col = c(2,3), legend = table((x-y > 0)), lty  = 1, bty = "n")
  legend("topright", lty = 1, col = "white", legend = c("***, p < 0.001", " **, p < 0.01", "  *, p < 0.05", "  -, p > 0.05"))
  text(x = 1.5, y = maxy - 1.65, pverdixy)
  lines(c(1, 2), c(maxy -1.75,maxy - 1.75))
  text(x = 2.5, y = maxy - 1.45, pverdiyz)
  lines(c(2, 3), c(maxy -1.55,maxy - 1.55))
  text(x = 2, y = maxy - 1.25, pverdixz)
  lines(c(1, 3), c(maxy -1.35,maxy - 1.35))
}


plotPerGenISamme <- function(x, y, z, IDx2, IDy2, IDx3, IDy3, tittel1, tittel2, tittel3){
  miny <- min(c(x,y,z))
  maxy <- max(c(x,y,z) + 2)

  par(mfrow = c(1,3))
  plotPerGenT1T2(x, y, z, tittel1, miny, maxy)
  plotPerGenT1T2(x[IDx2], y[IDy2], z, tittel2, miny, maxy)
  plotPerGenT1T2(x[IDx3], y[IDy3], z, tittel3, miny, maxy)
}


```


```{r, include = F}
#tar ut de med antall sykedager før T1 som her fler enn 10:
#table(dat1$Sykedager.før.T1 < 11)
```

```{r, include=F}
minst5prosent <- function(x){
  y <- x[x>min(x)]
  return(length(y)/length(x) > 0.05)
}
```

```{r, include = F, fig.width= 14, fig.height=18}
#dat1 <- dat1[dat1$Sykedager.før.T1 < 11,]
dat1.gener <- dat1[,gener]
dim(dat1.gener)
GenerMed <- apply(dat1.gener, 2, minst5prosent)
dat1.gener <- dat1.gener[,GenerMed]
dim(dat1.gener)


#dat2 <- dat2[dat2$Sykedager.før.T1 < 11, ]
dat2.gener <- dat2[,gener]
dat2.gener <- dat2.gener[,GenerMed]
#which(!apply(dat2.gener, 2, minst5prosent))
# apply(dat2.gener[,which(!apply(dat2.gener, 2, minst5prosent))], 2, sd)

dKont.gener <- dKont[,gener]
dKont.gener <- dKont.gener[,GenerMed]
#which(!apply(dKont.gener, 2, minst5prosent))

# apply(dKont.gener[,which(!apply(dKont.gener, 2, minst5prosent))], 2, sd)

# par(mfrow = c(5,3))
# for(i in 1:nrow(dat1.gener)){
#   hist(dat1.gener[,i], main = colnames(dat1.gener[i]))
# }


PasientT1T2 <- as.character(dat1$Patient_ID[dat1$Patient_ID %in% dat2$Patient_ID])
rownames(dat1.gener) <- as.character(dat1$Patient_ID)
rownames(dat2.gener) <- as.character(dat2$Patient_ID)
rownames(dat1) <- as.character(dat1$Patient_ID)
rownames(dat2) <- as.character(dat2$Patient_ID)

```


```{r, echo=FALSE}
vulcanoPlot <- function(X, Y, logTrans = TRUE, pairwise = FALSE, main = ""){
  
  n <- ncol(X)
  if(!ncol(Y) == n){
    knitr::kable("need as equaly many coloums in the two matrixes of log data")
  }
  if(logTrans == FALSE){
    X <- log2(X)
    Y <- log2(Y)
  } 
  #browser()
 
#   log2FoldCh <- apply(X, 2, median) - apply(Y, 2, median)
#   pvalue <- rep(NA, n)
#   for(i in 1:n){
#     pvalue[i] <- wilcox.test(X[,i], Y[,i], paired = pairwise)$p.value
#   }
#   adj.p.val <- p.adjust(pvalue, method = "fdr")
#   
#   diffexp <- NA
#   de <- data.frame(log2FoldChange = log2FoldCh, adj.p.value = adj.p.val)
#   de$logadj <- -log10(de$adj.p.value)
#   
#   de$diffexpressed <- "NO"
#   # if log2Foldchange > 0.6 and pvalue < 0.05, set as "UP" 
#   de$diffexpressed[de$log2FoldChange > 0.6 & de$adj.p.value < 0.05] <- "UP"
#   # if log2Foldchange < -0.6 and pvalue < 0.05, set as "DOWN"
#   de$diffexpressed[de$log2FoldChange < -0.6 & de$adj.p.value < 0.05] <- "DOWN"
#   
#   de$delabel <- NA
#   de$delabel[de$diffexpressed != "NO"] <- rownames(de)[de$diffexpressed != "NO"]
# 
#  
#   cols <- c("DOWN" = "blue", "NO" = "black", "UP" = "red")
#   
# 
#  aktuelleGener <- de[de$diffexpressed != "NO",]
# aktuelleGener$adj.p.value <- round(aktuelleGener$adj.p.value, 4)
#     gg <- ggplot(data=de, aes(x = log2FoldChange, y = -log10(adj.p.value), col = diffexpressed, label = delabel)) +   
#  #    gg <- ggplot(data=de, aes(x = log2FoldChange, y = adj.p.value, col = diffexpressed, label = delabel)) +
#         geom_point() + 
#         theme_minimal() +
#         geom_text_repel() +
#       theme_classic(base_size = 15) +
#         scale_color_manual(values=cols) +
#         geom_vline(xintercept=c(-0.6, 0.6), col="red") +
#         geom_hline(yintercept=-log10(0.05), col="red") + ylim(0,16) + xlim(-4,4) +
#         ggtitle(main)# +
#       #  scale_y_log10() 
#     
#     
    
    
    
  log2FoldCh <- apply(X, 2, median) - apply(Y, 2, median)
  pvalue <- rep(NA, n)
  for(i in 1:n){
    pvalue[i] <- wilcox.test(X[,i], Y[,i], paired = pairwise)$p.value
  }
  adj.p.val <- p.adjust(pvalue, method = "fdr")
  
  diffexp <- NA
  de <- data.frame(FoldChange = 2^log2FoldCh, adj.p.value = adj.p.val)
  de$logadj <- -log10(de$adj.p.value)
  
  de$diffexpressed <- "NO"
  # if log2Foldchange > 0.6 and pvalue < 0.05, set as "UP" 
  de$diffexpressed[de$FoldChange > 2^0.6 & de$adj.p.value < 0.05] <- "UP"
  # if log2Foldchange < -0.6 and pvalue < 0.05, set as "DOWN"
  de$diffexpressed[de$FoldChange < 2^(-0.6) & de$adj.p.value < 0.05] <- "DOWN"
  
  de$delabel <- NA
  de$delabel[de$diffexpressed != "NO"] <- rownames(de)[de$diffexpressed != "NO"]

 
  cols <- c("DOWN" = "blue", "NO" = "black", "UP" = "red")
  

 aktuelleGener <- de[de$diffexpressed != "NO",]
aktuelleGener$adj.p.value <- round(aktuelleGener$adj.p.value, 4)
    gg <- ggplot(data=de, aes(x = FoldChange, y = -log10(adj.p.value), col = diffexpressed, label = delabel)) +   
 #    gg <- ggplot(data=de, aes(x = log2FoldChange, y = adj.p.value, col = diffexpressed, label = delabel)) +
        geom_point() + 
        theme_minimal() +
        geom_text_repel( max.overlaps = Inf, size = 5) +
      theme_classic(base_size = 25) +
        scale_color_manual(values=cols) +
      #   geom_vline(xintercept=2^c(-0.6, 0.6), col="red") +
      #   geom_hline(yintercept=-log10(0.05), col="red") + 
       ylim(0,16) + xlim(-4,4) +
        ggtitle(main) +
       coord_trans(x="log2", xlim = c(0.03,16)) +
 #   scale_x_continuous(breaks = c(0.063, 0.125, 0.25, 0.5, 1, 2, 4, 8, 16)) + 
    scale_x_continuous(breaks = c(0.063, 0.125, 0.25, 0.5, 1, 2, 4, 8, 16)) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
      #  scale_y_log10() 

  return(list(gg = gg, aktuelleGener = aktuelleGener, d = de))
}

# slettt
# gg <- ggplot(d2, aes(x= IRR, y = -log10(adj.p.value), col = diff, label = kort_navn)) + 
#     geom_point() + xlab("Incident rate ratio") +
#     theme_minimal() +
#     geom_text_repel() +
#     theme_classic(base_size = 15) +
#     scale_color_manual(values=cols) +
#     geom_vline(xintercept=c(2^-0.6, 2^0.6), col="red") +
#     ggtitle(main) + 
#     theme(legend.position = "none") + 
#     ylim(0,ymax) +  coord_trans(x="log2", xlim = c(0.03,30)) +
#     scale_x_continuous(breaks = c(0.063, 0.125, 0.25, 0.5, 1, 2, 4, 8, 16)) + 
#     theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 
```


```{r, echo=FALSE, warning= FALSE}
# all data

## paired time 1 and time 2
# fellesGener <- colnames(dat1.gener)[colnames(dat1.gener) %in% colnames(dat2.gener)]
# fellesID <- dat1$Pasientnr[dat1$Pasientnr %in% dat2$Pasientnr]
# d1 <- dat1[dat1$Pasientnr %in% fellesID,]
# d2 <- dat2[dat2$Pasientnr %in% fellesID,]
# 
# d1 <- d1[order(d1$Pasientnr), fellesGener]
# d2 <- d2[order(d2$Pasientnr), fellesGener]
# 
# v <- vulcanoPlot(d1, d2[,fellesGener], main = "T1 vs T2, paired", pairwise = TRUE)
# v$gg
# 
# tiff(fs::path(utSti, "vulcanoplot T1_T2 paired.tiff"), width = 500, height = 700)
#  v$gg + theme(legend.position = "none")
# dev.off()
# gT1T2parret <-  v$gg + theme(legend.position = "none")
# OppNedT1T2_parret <- v$aktuelleGener
# knitr::kable(OppNedT1T2_parret)
```



```{r, echo=FALSE, warning= FALSE}
## time 1 and controls 
# fellesGener <- colnames(dat1.gener)[colnames(dat1.gener) %in% colnames(dKont.gener)]
# 
# v <- vulcanoPlot(dat1.gener[,fellesGener], dKont.gener[,fellesGener], main = "T1 vs controls")
# v$gg
# OppNedT1Kont <- v$aktuelleGener
# knitr::kable(OppNedT1Kont)
# 
# tiff(fs::path(utSti, "vulcanoplot T1_control.tiff"), width = 500, height = 700)
#  v$gg + theme(legend.position = "none")
# dev.off()
# gT1C <-  v$gg + theme(legend.position = "none")


```


```{r, echo=FALSE, warning= FALSE}
## time 2 and controls
# fellesGener <- colnames(dat2.gener)[colnames(dat2.gener) %in% colnames(dKont.gener)]
# 
# v <- vulcanoPlot(dat2.gener[,fellesGener], dKont.gener[,fellesGener], main = "T2 vs controls")
# v$gg
# OppNedT2Kont <- v$aktuelleGener
# knitr::kable(OppNedT2Kont)
# 
# tiff(fs::path(utSti, "vulcanoplot T2_control.tiff"), width = 500, height = 700)
#  v$gg + theme(legend.position = "none")
# dev.off()
# 
# gT2C <-  v$gg + theme(legend.position = "none")

```



```{r, echo=FALSE, warning= FALSE}
#install.packages("ggVennDiagram")
library(ggVennDiagram)
# x <- list("T1vsC" = OppNedT1Kont$delabel,
#           "T1vsT2" = OppNedT1T2_parret$delabel,
#           "T2vsC" = OppNedT2Kont$delabel)
# 
# gvenn <- ggVennDiagram(x) + scale_color_brewer(palette = "Paired") + scale_x_continuous(expand = expansion(mult = .2))
# gvenn
# tiff(fs::path(utSti, "venn_T1_T2_C.tiff"), width = 700, height = 700)
# gvenn
# dev.off()
```



```{r, echo=FALSE, warning= FALSE}
library(ggpubr)

# tiff(fs::path(utSti, "Alle_venn_T1_T2_C.tiff"), width = 1500, height = 700)
# ggarrange(gT1T2parret, gT1C, gT2C, gvenn , ncol = 4, nrow = 1)
# dev.off()
```




## severe time 1 and moderate time 1 

```{r, echo=FALSE, warning= FALSE}
fellesGener <- colnames(dat1.gener)
table(dat1$Status)

v <- vulcanoPlot(dat1.gener[dat1$Status == "Severe", fellesGener], dat1.gener[dat1$Status == "Moderate", fellesGener], main = "Severe T1 vs moderate T1")
v$gg
OppNedT1 <- v$aktuelleGener
knitr::kable(OppNedT1)

library(vegan)
linestack(log2(v$x[!is.na(v$label)]), v$label[!is.na(v$label)], axis = T)

tiff(fs::path(utSti, "vulcanoplot ST1vsMT1.tiff"), width = 500, height = 700)
 v$gg + theme(legend.position = "none")
dev.off()

gST1vsMT1 <-  v$gg + theme(legend.position = "none")
```






## severe time 1 and control

```{r, echo=FALSE, warning= FALSE}
fellesGener <- colnames(dat1.gener)[colnames(dat1.gener) %in% colnames(dKont.gener)]

v <- vulcanoPlot(dat1.gener[dat1$Status == "Severe",fellesGener], dKont.gener[,fellesGener], main = "Severe T1 vs controls")
v$gg
OppNedST1Kont <- v$aktuelleGener
knitr::kable(OppNedST1Kont)

d <- v$d
d <- d[!d$diffexpressed == "NO",]
d$col <- "red"
d$col[d$diffexpressed == "DOWN"] <- "blue"
d <- d[order(d$FoldChange),]
d$stjerner <- lapply(d$adj.p.value, pVerdiStjerne)

tiff(fs::path(utSti, "log2foldchange ST1_control.tiff"), width = 200, height = 700)
par(mar = c(5, 4, 4, 2) + 0.1)
linestack(log2(d$FoldChange), d$delabel, axis = T, air = 1.1, col = d$col, main = "Severe T1 vs controls", ylim = c(-4,4))
dev.off()

d$lab2 <- paste0(d$delabel, " (", d$stjerner, ")")

tiff(fs::path(utSti, "log2foldchange ST1_control_med stjerne.tiff"), width = 240, height = 700)
par(mar = c(5, 4, 4, 2) + 0.1)
linestack(log2(d$FoldChange), d$lab2, axis = T, air = 1.1, col = d$col, main = "Severe T1 vs controls", ylim = c(-4,4))
dev.off()


tiff(fs::path(utSti, "foldchange ST1_control.tiff"), width = 200, height = 700)
linestack(log2(d$FoldChange), d$delabel, axis = F, air = 1.1, col = d$col, main = "Severe T1 vs controls", ylim = c(-4,4)) 
axis(side = 2, at = (-4):4, label = 2^((-4):4))
dev.off()


tiff(fs::path(utSti, "foldchange ST1_control_med stjerne.tiff"), width = 240, height = 700)
linestack(log2(d$FoldChange), d$lab2, axis = F, air = 1.1, col = d$col, main = "Severe T1 vs controls", ylim = c(-4,4)) 
axis(side = 2, at = (-4):4, label = 2^((-4):4))
dev.off()




tiff(fs::path(utSti, "vulcanoplot ST1_control.tiff"), width = 500, height = 700)
 v$gg + theme(legend.position = "none")
dev.off()

gST1C <-  v$gg + theme(legend.position = "none")
```



## moderate time 1 and constrols 

```{r, echo=FALSE, warning= FALSE}
fellesGener <- colnames(dat1.gener)[colnames(dat1.gener) %in% colnames(dKont.gener)]

v <- vulcanoPlot(dat1.gener[dat1$Status == "Moderate", fellesGener], dKont.gener[,fellesGener], main = "Moderate T1 vs controls")
v$gg
OppNedMT1Kont <- v$aktuelleGener
knitr::kable(OppNedMT1Kont)

tiff(fs::path(utSti, "vulcanoplot MT1_control.tiff"), width = 500, height = 700)
 v$gg + theme(legend.position = "none")
dev.off()



d <- v$d
d <- d[!d$diffexpressed == "NO",]
d$col <- "red"
d$col[d$diffexpressed == "DOWN"] <- "blue"
d <- d[order(d$FoldChange),]
d$stjerner <- lapply(d$adj.p.value, pVerdiStjerne)

tiff(fs::path(utSti, "log2foldchange MT1_control.tiff"), width = 200, height = 700)
par(mar = c(5, 4, 4, 2) + 0.1)
linestack(log2(d$FoldChange), d$delabel, axis = T, air = 1.1, col = d$col, main = "Moderate T1 vs controls", ylim = c(-4,4))
dev.off()

d$lab2 <- paste0(d$delabel, " (", d$stjerner, ")")

tiff(fs::path(utSti, "log2foldchange MT1_control_med stjerne.tiff"), width = 240, height = 700)
par(mar = c(5, 4, 4, 2) + 0.1)
linestack(log2(d$FoldChange), d$lab2, axis = T, air = 1.1, col = d$col, main = "Moderate T1 vs controls", ylim = c(-4,4))
dev.off()


tiff(fs::path(utSti, "foldchange MT1_control.tiff"), width = 200, height = 700)
linestack(log2(d$FoldChange), d$delabel, axis = F, air = 1.1, col = d$col, main = "Moderate T1 vs controls", ylim = c(-4,4)) 
axis(side = 2, at = (-4):4, label = 2^((-4):4))
dev.off()


tiff(fs::path(utSti, "foldchange MT1_control_med stjerne.tiff"), width = 240, height = 700)
linestack(log2(d$FoldChange), d$lab2, axis = F, air = 1.1, col = d$col, main = "Moderate T1 vs controls", ylim = c(-4,4)) 
axis(side = 2, at = (-4):4, label = 2^((-4):4))
dev.off()






gMT1C <-  v$gg + theme(legend.position = "none")
```





```{r, echo=FALSE, warning= FALSE}
#install.packages("ggVennDiagram")
library(ggVennDiagram)
x <- list("ST1vsC" = OppNedST1Kont$delabel,
          "ST1vsMT1" = OppNedT1$delabel,
          "MT1vsC" = OppNedMT1Kont$delabel)

gvenn <- ggVennDiagram(x) + scale_color_brewer(palette = "Paired") + scale_x_continuous(expand = expansion(mult = .2))
gvenn
tiff(fs::path(utSti, "venn_ST1_MT1_C.tiff"))
gvenn
dev.off()
```




```{r, echo=FALSE, warning= FALSE}
library(ggpubr)

tiff(fs::path(utSti, "Alle_venn_ST1_MT1_C.tiff"), width = 1500, height = 700)
ggarrange(gST1vsMT1, gST1C, gMT1C, gvenn , ncol = 4, nrow = 1)
dev.off()



```

## severe time 2 and moderate time 2 

```{r, echo=FALSE, warning= FALSE}
fellesGener <- colnames(dat2.gener)
table(dat2$Status)

v <- vulcanoPlot(dat2.gener[dat2$Status == "Severe", fellesGener], dat2.gener[dat2$Status == "Moderate", fellesGener], main = "Severe T2 vs moderate T2")
v$gg
OppNedT2 <- v$aktuelleGener
knitr::kable(OppNedT2)

fellesGener <- colnames(dat2.gener)[colnames(dat2.gener) %in% colnames(dKont.gener)]


tiff(fs::path(utSti, "vulcanoplot ST2_MT2.tiff"), width = 500, height = 700)
 v$gg + theme(legend.position = "none")
dev.off()

gST2vsMT2 <-  v$gg + theme(legend.position = "none")





d <- v$d
d <- d[!d$diffexpressed == "NO",]
d$col <- "red"
d$col[d$diffexpressed == "DOWN"] <- "blue"
d <- d[order(d$FoldChange),]
d$stjerner <- lapply(d$adj.p.value, pVerdiStjerne)

tiff(fs::path(utSti, "log2foldchange ST2_MT2.tiff"), width = 200, height = 700)
par(mar = c(5, 4, 4, 2) + 0.1)
linestack(log2(d$FoldChange), d$delabel, axis = T, air = 1.1, col = d$col, main = "Severe T2 vs Moderate T2", ylim = c(-4,4))
dev.off()

d$lab2 <- paste0(d$delabel, " (", d$stjerner, ")")

tiff(fs::path(utSti, "log2foldchange ST2_MT2_med stjerne.tiff"), width = 240, height = 700)
par(mar = c(5, 4, 4, 2) + 0.1)
linestack(log2(d$FoldChange), d$lab2, axis = T, air = 1.1, col = d$col, main = "Severe T2 vs Moderate T2", ylim = c(-4,4))
dev.off()


tiff(fs::path(utSti, "foldchange ST2_MT2.tiff"), width = 200, height = 700)
linestack(log2(d$FoldChange), d$delabel, axis = F, air = 1.1, col = d$col, main = "Severe T2 vs Moderate T2", ylim = c(-4,4)) 
axis(side = 2, at = (-4):4, label = 2^((-4):4))
dev.off()


tiff(fs::path(utSti, "foldchange ST2_MT2_med stjerne.tiff"), width = 240, height = 700)
linestack(log2(d$FoldChange), d$lab2, axis = F, air = 1.1, col = d$col, main = "Severe T2 vs Moderate T2", ylim = c(-4,4)) 
axis(side = 2, at = (-4):4, label = 2^((-4):4))
dev.off()





```


## severe time 2 and control

```{r, echo=FALSE, warning= FALSE}
fellesGener <- colnames(dat2.gener)[colnames(dat2.gener) %in% colnames(dKont.gener)]

v <- vulcanoPlot(dat2.gener[dat2$Status == "Severe",fellesGener], dKont.gener[,fellesGener], main = "Severe T2 vs controls")
v$gg + theme(legend.position = "none")
OppNedST2Kont <- v$aktuelleGener
knitr::kable(OppNedST2Kont)

tiff(fs::path(utSti, "vulcanoplot ST2_control.tiff"), width = 500, height = 700)
 v$gg + theme(legend.position = "none")
dev.off()

gST2C <-  v$gg + theme(legend.position = "none")





d <- v$d
d <- d[!d$diffexpressed == "NO",]
d$col <- "red"
d$col[d$diffexpressed == "DOWN"] <- "blue"
d <- d[order(d$FoldChange),]
d$stjerner <- lapply(d$adj.p.value, pVerdiStjerne)

tiff(fs::path(utSti, "log2foldchange ST2_Controls.tiff"), width = 200, height = 700)
par(mar = c(5, 4, 4, 2) + 0.1)
linestack(log2(d$FoldChange), d$delabel, axis = T, air = 1.1, col = d$col, main = "Severe T2 vs Controls", ylim = c(-4,4))
dev.off()

d$lab2 <- paste0(d$delabel, " (", d$stjerner, ")")

tiff(fs::path(utSti, "log2foldchange ST2_Controls_med stjerne.tiff"), width = 240, height = 700)
par(mar = c(5, 4, 4, 2) + 0.1)
linestack(log2(d$FoldChange), d$lab2, axis = T, air = 1.1, col = d$col, main = "Severe T2 vs Controls", ylim = c(-4,4))
dev.off()


tiff(fs::path(utSti, "foldchange ST2_Controls.tiff"), width = 200, height = 700)
linestack(log2(d$FoldChange), d$delabel, axis = F, air = 1.1, col = d$col, main = "Severe T2 vs Controls", ylim = c(-4,4)) 
axis(side = 2, at = (-4):4, label = 2^((-4):4))
dev.off()


tiff(fs::path(utSti, "foldchange ST2_Controls_med stjerne.tiff"), width = 240, height = 700)
linestack(log2(d$FoldChange), d$lab2, axis = F, air = 1.1, col = d$col, main = "Severe T2 vs Controls", ylim = c(-4,4)) 
axis(side = 2, at = (-4):4, label = 2^((-4):4))
dev.off()


```



## moderate time 2 and constrols 

```{r, echo=FALSE, warning= FALSE}
fellesGener <- colnames(dat2.gener)[colnames(dat2.gener) %in% colnames(dKont.gener)]

v <- vulcanoPlot(dat2.gener[dat2$Status == "Moderate", fellesGener], dKont.gener[,fellesGener], main = "Moderate T2 vs controls")
v$gg
OppNedMT2Kont <- v$aktuelleGener
knitr::kable(OppNedMT2Kont)

tiff(fs::path(utSti, "vulcanoplot MT2_control.tiff"), width = 500, height = 700)
 v$gg + theme(legend.position = "none")
dev.off()

gMT2C <-  v$gg + theme(legend.position = "none")



d <- v$d
d <- d[!d$diffexpressed == "NO",]
d$col <- "red"
d$col[d$diffexpressed == "DOWN"] <- "blue"
d <- d[order(d$FoldChange),]
d$stjerner <- lapply(d$adj.p.value, pVerdiStjerne)

tiff(fs::path(utSti, "log2foldchange MT2_Controls.tiff"), width = 200, height = 700)
par(mar = c(5, 4, 4, 2) + 0.1)
linestack(log2(d$FoldChange), d$delabel, axis = T, air = 1.1, col = d$col, main = "Moderate T2 vs Controls", ylim = c(-4,4))
dev.off()

d$lab2 <- paste0(d$delabel, " (", d$stjerner, ")")

tiff(fs::path(utSti, "log2foldchange MT2_Controls_med stjerne.tiff"), width = 240, height = 700)
par(mar = c(5, 4, 4, 2) + 0.1)
linestack(log2(d$FoldChange), d$lab2, axis = T, air = 1.1, col = d$col, main = "Moderate T2 vs Controls", ylim = c(-4,4))
dev.off()


tiff(fs::path(utSti, "foldchange MT2_Controls.tiff"), width = 200, height = 700)
linestack(log2(d$FoldChange), d$delabel, axis = F, air = 1.1, col = d$col, main = "Moderate T2 vs Controls", ylim = c(-4,4)) 
axis(side = 2, at = (-4):4, label = 2^((-4):4))
dev.off()


tiff(fs::path(utSti, "foldchange MT2_Controls_med stjerne.tiff"), width = 240, height = 700)
linestack(log2(d$FoldChange), d$lab2, axis = F, air = 1.1, col = d$col, main = "Moderate T2 vs Controls", ylim = c(-4,4)) 
axis(side = 2, at = (-4):4, label = 2^((-4):4))
dev.off()



```











```{r, echo=FALSE, warning= FALSE}
#install.packages("ggVennDiagram")
library(ggVennDiagram)
x <- list("ST2vsC" = OppNedST2Kont$delabel,
          "ST2vsMT2" = OppNedT2$delabel,
          "MT2vsC" = OppNedMT2Kont$delabel)

gvenn <- ggVennDiagram(x) + scale_color_brewer(palette = "Paired") + scale_x_continuous(expand = expansion(mult = .2))
gvenn
tiff(fs::path(utSti, "venn_ST2_MT2_C.tiff"))
gvenn
dev.off()
```




```{r, echo=FALSE, warning= FALSE}

tiff(fs::path(utSti, "Alle_venn_ST2_MT2_C.tiff"), width = 1500, height = 700)
ggarrange(gST2vsMT2, gST2C, gMT2C, gvenn , ncol = 4, nrow = 1)
dev.off()
```



```{r, echo=FALSE, warning= FALSE}
# # KUN H3N2
# 
# dat1alle <- dat1
# dat2alle <- dat2
# dKontalle <- dKont
# dat1.gener.alle <- dat1.gener
# dat2.gener.alle <- dat2.gener
# 
# dat1.gener <- dat1.gener[dat1$Subgroup == "H3N2", ]
# dat2.gener <- dat2.gener[dat2$Subgroup == "H3N2", ]
# 
# dat1 <- dat1[dat1$Subgroup == "H3N2", ]
# dat2 <- dat2[dat2$Subgroup == "H3N2", ]
# 
# ```
# 
# 
# ## paired time 1 and time 2
# 
# ```{r, echo=FALSE, warning= FALSE}
# fellesGener <- colnames(dat1.gener)[colnames(dat1.gener) %in% colnames(dat2.gener)]
# fellesID <- dat1$Pasientnr[dat1$Pasientnr %in% dat2$Pasientnr]
# d1 <- dat1[dat1$Pasientnr %in% fellesID,]
# d2 <- dat2[dat2$Pasientnr %in% fellesID,]
# 
# d1 <- d1[order(d1$Pasientnr), fellesGener]
# d2 <- d2[order(d2$Pasientnr), fellesGener]
# 
# v <- vulcanoPlot(d1, d2[,fellesGener], main = "T1 vs T2, paired, H3N2", pairwise = TRUE)
# v$gg
# OppNedT1T2_parret <- v$aktuelleGener
# knitr::kable(OppNedT1T2_parret)
# ```
# 
# 
# ## time 1 and controls 
# 
# ```{r, echo=FALSE, warning= FALSE}
# fellesGener <- colnames(dat1.gener)[colnames(dat1.gener) %in% colnames(dKont.gener)]
# 
# v <- vulcanoPlot(dat1.gener[,fellesGener], dKont.gener[,fellesGener], main = "T1 vs controls, H3N2")
# v$gg
# OppNedT1Kont <- v$aktuelleGener
# knitr::kable(OppNedT1Kont)
# ```
# 
# ## time 2 and controls
# 
# ```{r, echo=FALSE, warning= FALSE}
# fellesGener <- colnames(dat2.gener)[colnames(dat2.gener) %in% colnames(dKont.gener)]
# 
# v <- vulcanoPlot(dat2.gener[,fellesGener], dKont.gener[,fellesGener], main = "T2 vs controls, H3N2")
# v$gg
# OppNedT2Kont <- v$aktuelleGener
# knitr::kable(OppNedT2Kont)
# ```
# 
# 
# ## severe time 1 and moderate time 1 
# 
# ```{r, echo=FALSE, warning= FALSE}
# fellesGener <- colnames(dat1.gener)
# table(dat1$Status)
# 
# v <- vulcanoPlot(dat1.gener[dat1$Status == "Severe", fellesGener], dat1.gener[dat1$Status == "Moderate", fellesGener], main = "Severe T1 vs moderate T1, H3N2")
# v$gg
# OppNed <- v$aktuelleGener
# knitr::kable(OppNed)
# ```
# 
# 
# 
# ## severe time 2 and moderate time 2 
# 
# ```{r, echo=FALSE, warning= FALSE}
# fellesGener <- colnames(dat2.gener)
# table(dat2$Status)
# 
# v <- vulcanoPlot(dat2.gener[dat2$Status == "Severe", fellesGener], dat2.gener[dat2$Status == "Moderate", fellesGener], main = "Severe T2 vs moderate T2, H3N2")
# v$gg
# OppNed <- v$aktuelleGener
# knitr::kable(OppNed)
# 
# ```
# 
# 
# 
# 
# ## severe time 1 and control
# 
# ```{r, echo=FALSE, warning= FALSE}
# fellesGener <- colnames(dat1.gener)[colnames(dat1.gener) %in% colnames(dKont.gener)]
# 
# v <- vulcanoPlot(dat1.gener[dat1$Status == "Severe",fellesGener], dKont.gener[,fellesGener], main = "Severe T1 vs controls")
# v$gg
# OppNed <- v$aktuelleGener
# knitr::kable(OppNed)
# ```
# 
# 
# 
# ## moderate time 1 and constrols 
# 
# ```{r, echo=FALSE, warning= FALSE}
# fellesGener <- colnames(dat1.gener)[colnames(dat1.gener) %in% colnames(dKont.gener)]
# 
# v <- vulcanoPlot(dat1.gener[dat1$Status == "Moderate", fellesGener], dKont.gener[,fellesGener], main = "Moderate T1 vs controls, H3N2")
# v$gg
# OppNed <- v$aktuelleGener
# knitr::kable(OppNed)
# ```
# 
# 
# 
# ## severe time 2 and control
# 
# ```{r, echo=FALSE, warning= FALSE}
# fellesGener <- colnames(dat2.gener)[colnames(dat2.gener) %in% colnames(dKont.gener)]
# 
# v <- vulcanoPlot(dat2.gener[dat2$Status == "Severe",fellesGener], dKont.gener[,fellesGener], main = "Severe T2 vs controls, H3N2")
# v$gg
# OppNed <- v$aktuelleGener
# knitr::kable(OppNed)
# ```
# 
# 
# 
# ## moderate time 2 and constrols 
# 
# ```{r, echo=FALSE, warning= FALSE}
# fellesGener <- colnames(dat2.gener)[colnames(dat2.gener) %in% colnames(dKont.gener)]
# 
# v <- vulcanoPlot(dat2.gener[dat2$Status == "Moderate", fellesGener], dKont.gener[,fellesGener], main = "Moderate T2 vs controls, H3N2")
# v$gg
# OppNed <- v$aktuelleGener
# knitr::kable(OppNed)
```